Circular Regression Trees and Forests with an Application to Probabilistic Wind Direction Forecasting

### 2020-07-05 pkg/circtree/inst/wind_paper/draft/JRSSC/rebuttal.tex

\begin{re}
This is a well-written paper on using circular regression trees for analyzing a
specific type of circular data. The probabilistic approach is well-founded. The
experiments on wind direction data are well organized and the results are
thoroughly analyzed.
\end{re}

\begin{re}
The authors could address model order determination problem using the
Bayes Information Criterion (Schwartz, 1978, Annals of Statistics).  
\end{re}

First of all we want to thank you for your positive feedback and valuable input.

A very good remark, indeed split variable selection based on BIC creates an
alternative possible tree-building strategy. Per default when building a
circular distributional tree, the package \textbf{circtree} uses
\emph{p}-values from association tests both for selecting the split variable
and for a so-called pre-pruning strategy which stops growing the tree as soon
as no significant associations can be found in a given subgroup. However, as
discussed by Zeileis~\emph{et al.}~(2008), an alternative strategy is to use
information criteria such as AIC (Akaike information criterion) or BIC (Bayes
information criterion) for cost-complexity post-pruning, where a large tree is
first built and then pruned back. We now discuss this aspect in the Appendix~A
of the manuscript and also refer to the submitted manuscript by
Schlosser~\emph{et al.}~2019 which investigates the performance of different
testing and pruning strategies employed in unbiased recursive partitioning
algorithms.

### 2020-07-05 pkg/circtree/inst/wind_paper/draft/JRSSC/rebuttal.tex

\begin{re}
This is a well-written paper on using circular regression trees for analyzing a
specific type of circular data. The probabilistic approach is well-founded. The
experiments on wind direction data are well organized and the results are
thoroughly analyzed.
\end{re}

\begin{re}
The authors could address model order determination problem using the
Bayes Information Criterion (Schwartz, 1978, Annals of Statistics).  
\end{re}

First of all we want to thank you for your positive feedback and valuable input.

A very good remark, indeed split variable selection based on BIC creates an
alternative possible tree-building strategy. Per default when building a
circular distributional tree, the package \textbf{circtree} uses
\emph{p}-values from association tests both for selecting the split variable
and for a so-called pre-pruning strategy which stops growing the tree as soon
as no significant associations can be found in a given subgroup. However, as
discussed in Zeileis~\emph{et al.}~(2008), an alternative strategy is to use
information criteria such as AIC (Akaike information criterion) or BIC (Bayes
information criterion) for cost-complexity post-pruning, where a large tree is
first built and then pruned back. We now discuss post-pruning in the Appendix~A
of the manuscript and also refer to the submitted manuscript by
Schlosser~\emph{et al.}~2019, which investigates the performance of different
testing and pruning strategies employed in unbiased recursive partitioning
algorithms.

### 2020-07-05 pkg/circtree/inst/wind_paper/draft/draft.Rnw
abstract shortened by 3 words
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3047 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 323199985b09c387192f18eeb11dd3542f1216b6
rewrote discussion of Hara & Chellappa, please check
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3048 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 7fa73b6b321aa82eaaeb885da2b9c11bd7e22d07

\Abstract{
%SUBMITTED VERSION:
%While circular data occur in a wide range of scientific fields, the methodology
%for distributional modeling and probabilistic forecasting of circular response
%variables is rather limited. Most of the existing methods are built on the
%framework of generalized linear and additive models, which are often
%challenging to optimize and interpret. Therefore, building on previous ideas
%for trees modeling circular means, we suggest a distributional approach for
%regression trees and random forests yielding probabilistic forecasts based on
%the von~Mises distribution. The resulting tree-based models simplify the
%estimation process by using the available covariates for partitioning the data
%into sufficiently homogeneous subgroups so that a simple von~Mises distribution
%without further covariates can be fitted to the circular response in each
%subgroup. These circular regression trees are straightforward to interpret, can
%capture nonlinear effects and interactions, and automatically select the
%relevant covariates that are associated with either location and/or scale
%changes in the von~Mises distribution. Combining an ensemble of circular
%regression trees to a circular regression forest can regularize and smooth the
%covariate effects. The new methods are evaluated in a case study on
%probabilistic wind direction forecasting at two Austrian airports, considering
%other common approaches as a benchmark.
%
%NEW LISA AND MORITZ VERSION
%Circular data occur in a wide range of scientific fields, however the
%methodology for distributional modeling and probabilistic forecasting of
%circular response variables is rather limited. While most of the existing
%methods are built within the framework of generalized linear and additive
%models, which are often challenging to optimize and interpret, we propose a
%distributional approach for regression trees yielding probabilistic forecasts
%based on the von~Mises distribution. The resulting tree-based models use the
%available covariates for partitioning the data into sufficiently homogeneous
%subgroups, allowing for a simple circular distributional fit in each subgroup.
%These circular regression trees are straightforward to interpret, can capture
%nonlinear effects and interactions, and automatically select the relevant
%covariates that are associated with changes in any distribution parameter.
%Ensembles or forests of such circular trees can further improve the forecasts
%by regularizing and stabilizing the model. The newly developed methodology is
%illustrated in an application on short-term probabilistic wind direction
%forecasting for Innsbruck and Vienna Airport required to direct airplanes to a
%safe landing. The predictive skill of the novel approaches is benchmarked with
%an additive regression model plus a persistency and a climatological model,
%employing the circular continuous ranked probability score.
While circular data occur in a wide range of scientific fields, the methodology 
for distributional modeling and probabilistic forecasting of circular response
variables is rather limited. Most of the existing methods are built on the
framework of generalized linear and additive models, which are often challenging
to optimize and interpret. Specifically, capturing abrupt changes or
interactions is not straightforward but often relevant, e.g., for modeling wind
directions subject to different wind regimes. Additionally, automatic covariate
selection is desirable when many predictor variables are available, as often the
case in weather forecasting.

@@

This novel approach to circular regression trees and forests complements the
literature on tree-based circular modeling. \cite{Lund:2002} already introduced
a circular regression tree algorithm where binary splits are made based on an
angular distance measure capturing node homogeneity. However, this only models
changes in the conditional mean, whereas modeling the conditional variance or
full probabilistic distribution would also allow uncertainty in the forecast to
be estimated \citep{Gneiting:2008}.  In the field of image based object
direction estimation, \citet{Hara+Chellappa:2017} presented a
distribution-based tree-building algorithm by applying $k$-means clustering to
the target space where the number of clusters $k$ is adaptively determined by
the Bayes information criterion employing the von~Mises distribution. Despite
their distribution-based splitting approach to circular response variables,
they solely predict the mode of the distribution, thus, providing no full
probabilistic forecasts. Combining these previous ideas for circular trees with
the concept of distributional modeling, we introduce a fully distributional
tree approach that not only considers splits in all distribution parameters,
but also yields a full probabilistic predictive model as proposed by
\citet{Gneiting:2008}. In addition, as a natural extension, ensembles or
forests of circular regression trees are presented in this study, showing that
these can further improve predictive power.

>>>>

\Abstract{
%SUBMITTED VERSION:
%While circular data occur in a wide range of scientific fields, the methodology
%for distributional modeling and probabilistic forecasting of circular response
%variables is rather limited. Most of the existing methods are built on the
%framework of generalized linear and additive models, which are often
%challenging to optimize and interpret. Therefore, building on previous ideas
%for trees modeling circular means, we suggest a distributional approach for
%regression trees and random forests yielding probabilistic forecasts based on
%the von~Mises distribution. The resulting tree-based models simplify the
%estimation process by using the available covariates for partitioning the data
%into sufficiently homogeneous subgroups so that a simple von~Mises distribution
%without further covariates can be fitted to the circular response in each
%subgroup. These circular regression trees are straightforward to interpret, can
%capture nonlinear effects and interactions, and automatically select the
%relevant covariates that are associated with either location and/or scale
%changes in the von~Mises distribution. Combining an ensemble of circular
%regression trees to a circular regression forest can regularize and smooth the
%covariate effects. The new methods are evaluated in a case study on
%probabilistic wind direction forecasting at two Austrian airports, considering
%other common approaches as a benchmark.
%
%NEW LISA AND MORITZ VERSION
%Circular data occur in a wide range of scientific fields, however the
%methodology for distributional modeling and probabilistic forecasting of
%circular response variables is rather limited. While most of the existing
%methods are built within the framework of generalized linear and additive
%models, which are often challenging to optimize and interpret, we propose a
%distributional approach for regression trees yielding probabilistic forecasts
%based on the von~Mises distribution. The resulting tree-based models use the
%available covariates for partitioning the data into sufficiently homogeneous
%subgroups, allowing for a simple circular distributional fit in each subgroup.
%These circular regression trees are straightforward to interpret, can capture
%nonlinear effects and interactions, and automatically select the relevant
%covariates that are associated with changes in any distribution parameter.
%Ensembles or forests of such circular trees can further improve the forecasts
%by regularizing and stabilizing the model. The newly developed methodology is
%illustrated in an application on short-term probabilistic wind direction
%forecasting for Innsbruck and Vienna Airport required to direct airplanes to a
%safe landing. The predictive skill of the novel approaches is benchmarked with
%an additive regression model plus a persistency and a climatological model,
%employing the circular continuous ranked probability score.
While circular data occur in a wide range of scientific fields, the methodology 
for distributional modeling and probabilistic forecasting of circular response
variables is rather limited. Most of the existing methods are built on
generalized linear and additive models, which are often challenging
to optimize and interpret. Specifically, capturing abrupt changes or
interactions is not straightforward but often relevant, e.g., for modeling wind
directions subject to different wind regimes. Additionally, automatic covariate
selection is desirable when many predictor variables are available, as often the
case in weather forecasting.

@@

This novel approach to circular regression trees and forests complements the
literature on tree-based circular modeling. \cite{Lund:2002} already introduced
a circular regression tree algorithm where binary splits are made based on an
angular distance measure capturing node homogeneity. However, this only models
changes in the conditional mean, whereas modeling the conditional variance or
full probabilistic distribution would also allow uncertainty in the forecast to
be estimated \citep{Gneiting:2008}.  Moreover, \citet{Hara+Chellappa:2017}
introduced both regression trees and random forests for image-based object
direction estimation. These extend classical conditional-mean regression trees
\citep{Breiman+Friedman+Stone:1984} and corresponding random forests
\citep{Breiman:2001} to conditional mean models for cicular data. But rather
than considering only binary splits, they employed $k$-means clustering
to determine multi-way splits with the number $k$ adaptivaly determined
based on the Bayes information criterion with either a Gaussian or von~Mises
kernel. While a distributional aspect is considered in this step of the
algorithm, variances are always restricted to be shared across splits and, more
importantly, no probabilistic forecasts are obtained. Therefore, our proposed
circular regression trees and random forests take these ideas a step further:
A fully distributional approach is adopted for all aspects of the algorithms
from selection of the split variables, selection of the split points, and
ensemble aggregation of the predictions across trees in a forest. Therefore,
this yields a full probabilistic predictive model as proposed by
\citet{Gneiting:2008}.
